Constant Error Tolerance Computational Fuzzy Extractors

### 2014-01-30 constantRate/ConstCompFuzzExt.tex
preliminary proof of security of sampling cons

\section{A Constant Rate Computational Fuzzy Extractor}
\label{sec:construction}

@@

%\noindent We argue security assuming that the adversary knows the value of all blocks outside of $\mathcal{I}$.  
We show security of \consref{cons:first construction} for distributions satisfying \defref{def:block guessable}.
To show security it is sufficient to show that there exists a distribution $C'$ with conditional min-entropy and $\delta^{\mathcal{D}_{s_{sec}}}((C, P), (C', P))$ are indistinguishable.  Let $\mathcal{I}$ be the set of indices that exists in \defref{def:block guessable}, the distribution $C'$ is defined as a uniform codeword conditioned on the values of $C$ and $C'$ being equal on all indices outside of $\mathcal{I}$.  First note that $\Hav(C' |P) =\Hav(C' | C_{\mathcal{I}^c} = C'_{\mathcal{I}^c}) = |C| - |\mathcal{I}^c|$~(\clref{cl:many locations ent}).  It is left to show $\delta^{\mathcal{D}_{s_{sec}}}((C, P), (C', P))$.  
Define the distribution $X$ as follows:
\[X_i = 
\begin{cases}
W_i & C_i = 0\\
R_i & C_i = 1.
\end{cases}\]
The outline for the rest of the proof is as follows:
\begin{itemize}
\item Let $D$ be a distinguisher between $(C, P)$ and $(C', P)$, since $P$ is a collection of obfuscated programs there exists a simulator, $S$~(outputting a single bit), such that $D(C, P) \approx_c S^{\mathcal{O}_X(\cdot, \cdot)}(C)$~(\defref{def:obf}).
\item Show that even an unbounded $S$ with oracle access to $X_1,..., X_\ell$ cannot distinguish between $C$ and $C'$.  That is, $S^{\mathcal{O}_X(\cdot, \cdot)}(C) \approx S^{\mathcal{O}_X(\cdot, \cdot)}(C')$.
This is done in \lemref{lem:sim cannot distinguish}.
% the following steps:
%\begin{itemize}
%%\item Show that conditioned on the values of all other indices, $\mathcal{I}^c$, $C$ still has many possible values.  That is, $\Hav(C | C_{\mathcal{I}^c})$ is large~(\clref{cl:many locations ent}).
%%\item Assume that the simulator has knowledge of $C_{\mathcal{I}^c}$.
%\item For all codewords $c, c' \in C| C_{\mathcal{I}^c}$, the codewords $c, c'$ are indistinguishable given the simulator's view~(\lemref{lem:codewords in I close}).
%\item The two distributions $C$ and $C'$ are indistinguishable given the simulator's view~(\bnote{this needs to be written down.})
%\item The set of codewords $C| View(S) \wedge C_{\mathcal{I}^c}$ has conditional min-entropy~(and thus $C|View(S)$ has conditional min-entropy)~(\corref{cor:avg min after view}).  
%\item Any adversary given $C$ or a fresh $C'$ cannot statistically distinguish.
%\end{itemize}
\item By the security of obfuscation, $S^{\mathcal{O}}(C') \approx_c D(C', P)$~(\defref{def:obf}).
\end{itemize}

@@

\subsection{Characterizing unguessable block distributions}

\defref{def:block guessable} is an inherently adaptive definition and thus a little unwieldy for a distribution.  In this section, we provide a limited characterization of sources satisfy \defref{def:block guessable} and distributions that are learnable by equality queries.

The majority of the difficulty in characterizing the definition is that different blocks may be dependent so a equality query on block $i$ may shape the distribution of block $j$.  Thus, we begin with the case of independent blocks.

\begin{claim}
\label{cl:independent high ent}
Let $W = W_1,..., W_\ell$ be a source where blocks are independent.  Let $\alpha$ be a parameter.  Let $\mathcal{I}$ be the set of blocks such that $\Hoo(W_i ) =\alpha $.  Then $W$ is a $(q, \alpha - \log (q+1), \ell - |\mathcal{I}|)$-block unguessable distribution.  In particular, when $\alpha = \omega(\log n)$ and $q = \poly(n)$, then $W$ is a $(q, \omega(\log n), \ell - |\mathcal{I}|)$-block unguessable distribution.
\end{claim}
\begin{proof}
We ignore the blocks outside of $\mathcal{I}$ as queries on these blocks do not change the distribution of $W_i$ where $i\in \mathcal{I}$.  It suffices to show that for all $i\in \mathcal{I}, \Hav(W_i) = \alpha -\log q$.  
Consider a particular block $i$, we measure how an adversary's queries affect the entropy of this block~(we can ignore the other blocks as they are independent).  
Let $q_w$ be a query asking if the stored value is $a_w$ and its corresponding response.  
Let $Q_{w_1},A_{w_1},..., Q_{w_q}, A_{w_q}$ be the random variables representing the queries and answers for an  adversary $\mathcal{A}$ making $q$ queries.  We assume a deterministic adversary (the adversary is unbounded and thus can compute the best possible queries).  The only dependence between $W_i$ and this view of the adversary is in $A_1,..., A_q$ so we can consider these binary responses.  Our goal is to count the total number of possible responses $A_{w_1},..., A_{w_q}$.  There are two basic cases for $A_{w_1},..., A_{w_q}$: the all zeros string and the case where some query returns $1$.  If some query $Q_{w*}$ returns $1$ then all other $Q_{w_i} = Q_{w*}$ will return $1$ but no other query returns $1$.  These responses can be removed as the response is deterministic.  Let $A'$ represent the sequence with all duplicate queries removed.  This sequence has at most a single $1$.  Thus, the total number of possible responses is $q+1$~(including the duplicates does not change the number of possible responses).  Thus, we have the following,
\begin{align*}
\Hav(W_i | View(S)) &= \Hav(W_i| Q_{w_1}, A_{w_1},..., Q_{w_q}, A_{w_q})\\
&=\Hav(W_i | A_{w_1},..., A_{w_q})\\
&=\Hav(W_i |A') \\
&=\Hoo(W_i) - \log |A'|\\
&=\alpha - \log |A'|\\
&=\alpha - \log (q+1)
\end{align*}
Where the fourth line follows from the third by~\cite[Lemma 2.2]{DBLP:journals/siamcomp/DodisORS08}.
This argument holds for all $i$.  This completes the proof.
\end{proof}
In their work on computational fuzzy extractors, Fuller, Meng, and Reyzin~\cite{fuller2013computational} consider block-fixing sources, where each block is either uniform or a fixed symbol~(block fixing sources were introduced by Kamp and Zuckerman~\cite{KZ07}).  This is a subset of the distributions in \clref{cl:independent high ent}.

We now consider more complicated~(realistic) distributions where blocks are not independent.  We begin with distributions that are block unguessable.

Note the total entropy of a source in \clref{cl:each block from single seed} is $e$, so this is a family of distributions with total entropy $\omega(\log n)$ where \consref{cons:first construction} is secure.  This claims rests on there being no easy ``entry'' point to the distribution.  This is not always the case.  Indeed it is possible for some blocks to have very high entropy but lose all of it after equality queries.

@@

\subsection{Security of \consref{cons:sampling}}
In order to show security \consref{cons:sampling}, we show that with overwhelming probability each of the selected $v_i$ has entropy on its own.  This means that  $V_1,..., V_\ell$ is a block-unguessable distribution and security follows by \thref{thm:security of cons}.  Consider a source with entropy $\lambda$.  The expected number of selected blocks that have entropy at least $\alpha$ is $\frac{\gamma-\beta}{\gamma} \times \beta$.  A good sampler selects a within a constant fraction of the expected number of these blocks.

\begin{lemma}
\bnote{write a lemma about this}
\end{lemma}
%\begin{align*}
%\expe[\Hoo(V_i)] &=\expe[\text{entropy of each uniform }w_j] (\text{number of }w_j\text{ selected})\\
%&=\frac{\lambda}{\gamma} \times \eta= \frac{\lambda \eta}{\gamma}.
%\end{align*}
%This value is super-logarithmic when $\lambda = \omega (\frac{\gamma \log n}{\eta})$.  Note this is independent of how the entropy is distributed between blocks.  In particular, we make no assumptions about particular blocks being unguessable, only about the overall entropy rate.  This is a much more natural condition for a source than \defref{def:block guessable}.  
%To show security we need the entropy of all $v_i$ to be $\omega(\log n)$ with overwhelming probability.  This follows by an argument from the locally computable extractor literature.

\begin{theorem}
\label{th:samp sec}
Let $n$ be a security parameter and let $Z$ be some super-polynomially size alphabet.  Let $W = W_1,..., W_\gamma$ be a distribution where each $W_i$ is over $Z$.  Let 
\[
\Hoo(W) = \omega\left(\frac{\gamma \log n}{\eta}\right),
\] 
then for all $i$ and for all values of $j_{1,1},..., j_{\ell, \eta}$ the distribution $\delta(V_i | j_{1,1},..., j_{\ell, \eta} , V'_i ) <\epsilon_{sam}$ where $\Hoo(V_i') = \omega(\log n)$.\bnote{I think $\epsilon_{sam}$ will be a negligible for resultant entropy of $O(\beta \times \alpha \times (\mathcal{I}-\beta)/\mathcal{I})$.}
\end{theorem}

\begin{corollary}
\label{cor:samp unguess}
Let $n, Z, W$ be as in \thref{th:samp sec}, then for all $j_{1,1},..., j_{\ell, \eta}$, $V | j_{1,1},..., j_{\ell,\eta}$ is $\epsilon_{sam}$ close to a $(\poly(n), \alpha = \omega(\log n), 0)$-block unguessable distribution.  In particular, If $\mathcal{O}$ is an $\ell$-composable virtual black obfuscator with auxiliary inputs for point programs over $\zo^{\beta}$ that is $(s_{obf}, \epsilon_{obf})$-hard, then $H^{\hill}_{\epsilon', s_{obf}}(C | P) \geq |C|$ for $\epsilon' = 2\epsilon_{obf} + \ell(\epsilon_{sam} + 2^{-(\alpha-1)})$.
\end{corollary}
\begin{proof}
Direct result of \thref{th:samp sec}, \bnote{need a claim saying if all $W_i$ have entropy, we are block unguessable}, and \thref{thm:security of cons}.
\end{proof}

\bnote{something something about by sampling theory the prob of this being more than constant from expected value is negligible.}
\bnote{need to finish this}

@@

>>>>

\section{A Constant Error-Tolerance Computational Fuzzy Extractor}
\label{sec:construction}

@@

%\noindent We argue security assuming that the adversary knows the value of all blocks outside of $\mathcal{I}$.  
We show security of \consref{cons:first construction} for distributions satisfying \defref{def:block guessable}.  In \apref{sec:characterize} we provide discussion on the types of sources that satisfy \defref{def:block guessable}.
To show security it is sufficient to show that there exists a distribution $C'$ with conditional min-entropy and $\delta^{\mathcal{D}_{s_{sec}}}((C, P), (C', P))$ are indistinguishable.  Let $\mathcal{I}$ be the set of indices that exists in \defref{def:block guessable}, the distribution $C'$ is defined as a uniform codeword conditioned on the values of $C$ and $C'$ being equal on all indices outside of $\mathcal{I}$.  First note that $\Hav(C' |P) =\Hav(C' | C_{\mathcal{I}^c} = C'_{\mathcal{I}^c}) = |C| - |\mathcal{I}^c|$~(\clref{cl:many locations ent}).  It is left to show $\delta^{\mathcal{D}_{s_{sec}}}((C, P), (C', P))$.  
Define the distribution $X$ as follows:
\[X_i = 
\begin{cases}
W_i & C_i = 0\\
R_i & C_i = 1.
\end{cases}\]
The outline for the rest of the proof is as follows:
\begin{itemize}
\item Let $D$ be a distinguisher between $(C, P)$ and $(C', P)$, since $P$ is a collection of obfuscated programs there exists a simulator, $S$~(outputting a single bit), such that $D(C, P) \approx_c S^{\mathcal{O}_X(\cdot, \cdot)}(C)$~(\defref{def:obf}).
\item Show that even an unbounded $S$ with oracle access to $X_1,..., X_\ell$ cannot distinguish between $C$ and $C'$.  That is, $S^{\mathcal{O}_X(\cdot, \cdot)}(C) \approx S^{\mathcal{O}_X(\cdot, \cdot)}(C')$.
This is done in \lemref{lem:sim cannot distinguish}.
% the following steps:
%\begin{itemize}
%%\item Show that conditioned on the values of all other indices, $\mathcal{I}^c$, $C$ still has many possible values.  That is, $\Hav(C | C_{\mathcal{I}^c})$ is large~(\clref{cl:many locations ent}).
%%\item Assume that the simulator has knowledge of $C_{\mathcal{I}^c}$.
%\item For all codewords $c, c' \in C| C_{\mathcal{I}^c}$, the codewords $c, c'$ are indistinguishable given the simulator's view~(\lemref{lem:codewords in I close}).
%\item The two distributions $C$ and $C'$ are indistinguishable given the simulator's view~(\bnote{this needs to be written down.})
%\item The set of codewords $C| View(S) \wedge C_{\mathcal{I}^c}$ has conditional min-entropy~(and thus $C|View(S)$ has conditional min-entropy)~(\corref{cor:avg min after view}).  
%\item Any adversary given $C$ or a fresh $C'$ cannot statistically distinguish.
%\end{itemize}
\item By the security of obfuscation, $S^{\mathcal{O}}(C') \approx_c D(C', P)$~(\defref{def:obf}).
\end{itemize}

@@

@@

\subsection{Security of \consref{cons:sampling}}
In order to show security \consref{cons:sampling}, we show that with overwhelming probability each of the selected $v_i$ has super-logarithmic entropy on its own.  This means that  $V_1,..., V_\ell$ is a block-unguessable distribution~(by \clref{cl:all blocks entropy}) and security follows by \thref{thm:security of cons}.  Consider a source with entropy $\lambda$.  The expected number of selected blocks that have entropy at least $\alpha$ is $\frac{\gamma-\beta}{\gamma} \times \beta$.  A good sampler selects within a constant fraction of the expected number of these blocks.

\begin{lemma}
\label{lem:sampling works}
Let $W = W_1,..., W_\gamma$ be a $(\alpha = \Omega(1), \beta \leq \gamma /2-\Omega(1))$-unordered block source.  Let $\eta = \Omega(\log n)$. Consider some fixed $i$, the distribution $V_{i}, J_{i, 1},..., J_{i, \eta}$ is $(\ngl(n)+ 2^{-\Omega(n)})$-close to a distribution $V', U_{\beta\times \log \gamma}$ where for all $u\in U_{\beta \times \log \gamma}$ $V' | A=a$, $\Hoo(V' | A=a) \geq \Omega(\alpha\times (1-2\times \beta) \times \eta) = \omega(\log n) = \alpha'$.\bnote{need to come up with the right number of coins for $\sample$}
\end{lemma}
\begin{proof}
By~\cite[Lemma 9]{vadhan2003constructing}, noting that the truly random sampler is optimum in necessary samples and we take $\eta = \Omega(\log n)$ samples.  \bnote{need someone to check parameters, I am fairly sure of the quality of the statement}
\end{proof}
%\begin{align*}
%\expe[\Hoo(V_i)] &=\expe[\text{entropy of each uniform }w_j] (\text{number of }w_j\text{ selected})\\
%&=\frac{\lambda}{\gamma} \times \eta= \frac{\lambda \eta}{\gamma}.
%\end{align*}
%This value is super-logarithmic when $\lambda = \omega (\frac{\gamma \log n}{\eta})$.  Note this is independent of how the entropy is distributed between blocks.  In particular, we make no assumptions about particular blocks being unguessable, only about the overall entropy rate.  This is a much more natural condition for a source than \defref{def:block guessable}.  
%To show security we need the entropy of all $v_i$ to be $\omega(\log n)$ with overwhelming probability.  This follows by an argument from the locally computable extractor literature.

\begin{corollary}
\label{cor:samp sec}
Let $W = W_1,..., W_\gamma$ be a $(\alpha = \Omega(1), \beta \leq \gamma /2-\Omega(1))$-unordered block source.  Let $\eta = \Omega(\log n)$.  
The distribution $V, J$ is $\epsilon_{sam}(\ell (\ngl(n) +2^{-\Omega(n)})) = \ngl(n)$-close to a distribution $V', U_{\beta\times \ell \times \log \gamma}$ where for $u\in U_{\beta\times \ell\times \log \gamma}$ for all $i$, $\Hoo(V_i' | U_{\beta\times \ell \times \log \gamma}) \geq \alpha'$.
\end{corollary}
\begin{proof}
Union bound over the statistical distance in \lemref{lem:sampling works}.
\end{proof}

\begin{corollary}
\label{cor:v are unguessable}
Let $q = \poly(n)$, there exists some $\alpha'' = \omega(\log n)$ such that distribution $V' , U_{\beta \times \log \gamma}$ is a $(q, \alpha'', 0)$-block unguessable distribution.
\end{corollary}
\begin{proof}
Result of \clref{cl:all blocks entropy}.
\end{proof}

\begin{corollary}
\label{cor:samp unguess}
Let $n, Z, W, \epsilon_{sam}$ be as in \corref{cor:samp sec} and let $\mathcal{O}$ is an $\ell$-composable virtual black obfuscator with auxiliary inputs for point programs over $\zo^{\beta\times \log |Z|}$ that is $(s_{obf}, \epsilon_{obf})$-hard, then $H^{\hill}_{\epsilon', s_{obf}}(C | P) \geq |C|$ for $\epsilon' = (2\epsilon_{obf} + \epsilon_{sam}+ 2^{-(\alpha-1)}) = \ngl(n)$.
\end{corollary}
\begin{proof}
Result of \corref{cor:v are unguessable} and \thref{thm:security of cons}.
\end{proof}

@@

\section{Characterizing unguessable block distributions}
\label{sec:characterize}

\defref{def:block guessable} is an inherently adaptive definition and thus a little unwieldy for a distribution.  In this section, we partially characterize sources that satisfy \defref{def:block guessable}.  
The majority of the difficulty in characterizing \defref{def:block guessable} is that different blocks may be dependent so a equality query on block $i$ may shape the distribution of block $j$.  Thus, we begin with the case of independent blocks.

\begin{claim}
\label{cl:independent high ent}
Let $W = W_1,..., W_\ell$ be a source where blocks are independent.  Let $\alpha$ be a parameter.  Let $\mathcal{I}$ be the set of blocks such that $\Hoo(W_i ) =\alpha $.  Then $W$ is a $(q, \alpha - \log (q+1), \ell - |\mathcal{I}|)$-block unguessable distribution.  In particular, when $\alpha = \omega(\log n)$ and $q = \poly(n)$, then $W$ is a $(q, \omega(\log n), \ell - |\mathcal{I}|)$-block unguessable distribution.
\end{claim}
\begin{proof}
We ignore the blocks outside of $\mathcal{I}$ as queries on these blocks do not change the distribution of $W_i$ where $i\in \mathcal{I}$.  It suffices to show that for all $i\in \mathcal{I}, \Hav(W_i) = \alpha -\log q$.  
Consider a particular block $i$, we measure how an adversary's queries affect the entropy of this block~(we ignore all other blocks as they are independent).  
Let $q_w$ be a query asking if the stored value is $a_w$ and its corresponding response.  
Let $Q_{w_1},A_{w_1},..., Q_{w_q}, A_{w_q}$ be the random variables representing the queries and answers for an  adversary $\mathcal{A}$ making $q$ queries.  We assume a deterministic adversary~(we assume the best coins are in nonuniform advice).  The only dependence between $W_i$ and this view of the adversary is in $A_1,..., A_q$ so we can consider these binary responses.  Our goal is to count the total number of possible responses $A_{w_1},..., A_{w_q}$.  There are two basic cases for $A_{w_1},..., A_{w_q}$: the all zeros string and the case where some query returns $1$.  If some query $Q_{w*}$ returns $1$ then all other $Q_{w_i} = Q_{w*}$ will return $1$ but no other query returns $1$.  These responses can be removed as the response is deterministic.  Let $A'$ represent the sequence with all duplicate queries removed.  This sequence has at most a single $1$.  Thus, the total number of possible responses is $q+1$~(including the duplicates does not change the number of possible responses).  Thus, we have the following,
\begin{align*}
\Hav(W_i | View(S)) &= \Hav(W_i| Q_{w_1}, A_{w_1},..., Q_{w_q}, A_{w_q})\\
&=\Hav(W_i | A_{w_1},..., A_{w_q})\\
&=\Hav(W_i |A') \\
&=\Hoo(W_i) - \log |A'|\\
&=\alpha - \log |A'|\\
&=\alpha - \log (q+1)
\end{align*}
Where the fourth line follows from the third by~\cite[Lemma 2.2]{DBLP:journals/siamcomp/DodisORS08}.
This argument holds for all $i$.  This completes the proof.
\end{proof}
In their work on computational fuzzy extractors, Fuller, Meng, and Reyzin~\cite{fuller2013computational} consider block-fixing sources, where each block is either uniform or a fixed symbol~(block fixing sources were introduced by Kamp and Zuckerman~\cite{KZ07}).  \clref{cl:independent high ent} captures this type of distribution.  
We now consider more complicated distributions where blocks are not independent.

Note the total entropy of a source in \clref{cl:each block from single seed} is $e$, so this is a family of distributions with total entropy $\omega(\log n)$ where \consref{cons:first construction} is secure.  We can prove a similar claim in the case where $f$ is not injectiveall blocks do not a source~(determined by $f$).

\begin{claim}
\label{cl:all blocks entropy}
Let $W = W_1,..., W_\ell$ be a source.  Let $\alpha =\omega(\log n)$ and suppose that for all $i$ $\Hoo(W_i)\geq \alpha$, then for $q = \poly(n)$, $\alpha ' = \omega(\log n)$, $W$ is a $(q, \alpha', \ell)$-block unguessable distribution.
\end{claim}
\begin{proof}
Consider a particular block $W_i$, it suffices to show that $\Hav(W_i | Q_1,A_1,..., Q_q, A_q) \geq \omega(\log n)$.  By the definition of conditional min-entropy:
\begin{align*}
\Hav(W_i | Q_1, A_1,..., Q_q, A_q) &= -\log (\Pr[\text{some non zero response}] \max_{w\in W_i} \Pr[W_i | \text{non zero response}) \\&+ \Pr[\text{all zero responses}]\max_{w\in W_i} \Pr[W_i = w_i| \text{all zero responses}])\\
&\geq-\log (\Pr[\text{some non zero response}]\times 1 \\&+ \Pr[\text{all zero responses}]\max_{w\in W_i} \Pr[W_i = w_i | \text{all zero responses}]).
\end{align*}
It suffices to show that $\Pr[\text{some non zero response}] = \ngl(n)$ and that $\max_{w\in W_i} \Pr[W_i | \text{all zero responses}] = \ngl(n)$.  We show each of these is turn.

First note that clearly the 1st query returns $1$ with $2^{-\alpha}$ probability~(due to the min-entropy condition).  Conditioned on this response being $0$ the probability of all other events increases by at most negligible amount~(the statistical distance between the original distribution and the distribution conditioned on the first response being $0$ is $\ngl(n)$).  Thus, the total probability of a $1$ is at most:
\[
\Pr[\text{some non zero response}] \leq 2^{-\alpha} + (2^{-\alpha}+\ngl(n)) + ... +(2^{-\alpha}+q\times \ngl(n)) = \ngl(n).
\]
Now consider $\max_{w\in W_i} \Pr[W_i | \text{all zero responses}]$.  For each zero response, the probability of any $w_i$ increases by at most a negligible amount, this means that 
\[\max_{w\in W_i} \Pr[W_i | \text{all zero responses}] = 2^{-\alpha} + q\times \ngl(n) = \ngl(n).\]

Thus, we have the following:
\begin{align*}
\Hav(W_i | Q_1, A_1,..., Q_q, A_q) &\geq-\log (\Pr[\text{some non zero response}]\times 1 \\&+ \Pr[\text{all zero responses}]\max_{w\in W_i} \Pr[W_i = w_i | \text{all zero responses}])\\
& \geq -\log (\ngl(n) + (1-\ngl(n))\ngl(n)) = -\log (\ngl(n)) = \omega(\log (n)).
\end{align*}
This completes the proof.\bnote{someone check this proof!!!! also make it more precise and better}
\end{proof}

Claims~\ref{cl:each block from single seed} and~\ref{cl:all blocks entropy} rest on there being no easy ``entry'' point to the distribution.  This is not always the case.  Indeed it is possible for some blocks to have very high entropy but lose all of it after equality queries.
